---
title: "BIOL3207_Assignment2"
author: "Yusheng_Wang_u7457423"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
```
# Git Hub Repo

[My GitHub Repository](https://github.com/moaaom/YUSHENG_BIOL3207_ASSIGNMENT2.git)

# Data Cleansing

## Load Packages and Data

First, let's load some useful packages.
```{r loadpackages, message = F, results='hide'}
# load packages with library()
library(bookdown)
library(tidyverse)
library(metafor)
library(orchaRd)
library(patchwork)
library(flextable)
```

Then, let's read data.
```{r readdata, message = F}
# read raw data from Clark et al. (2020)
crdata <- read_csv ("./raw_data/OA_activitydat_20190302_BIOL3207.csv")

# read metadata from Clark et al. (2020)
cmrdata <- read_csv("./raw_data/clark_paper_data.csv")

# read meta-analysis dataset
mdata <- read_csv("./raw_data/ocean_meta_data.csv")
```

## Data Analyses for Clark *et al.* (2020)

First, let's remove rows that include NA values. Because they are useless that we cannot guess what the values actually are.
```{r rmna, message = F}
# check the completeness of values with complete.cases() and remove NA with filter()
ncrdata <- crdata %>% filter(., complete.cases(.))

# have a quick look
head(ncrdata)
```

Then, let's drop out irrelevant columns. This time, we only need the column `species`, `treatment` and `activity`.
```{r rmirre}
# select and keep the three columns
cdata <- tibble(species = ncrdata$species, treatment = ncrdata$treatment, activity = ncrdata$activity)

# have a quick look
head(cdata)
```

classes of columns `species` and `treatment` should be factors. Let's change them.
```{r refactors, message = F}
# change the last three column to factors
cdata$species <- as.factor(cdata$species)
cdata$treatment <- as.factor(cdata$treatment)

# have a quick look
head(cdata)
```

In this tibble, the columns, from left to right, are:
```{r colmeaning, message = F, eval = F}
1) Species name: acantho = Acanthochromis polyacanthus; Ambon = Pomacentrus amboinensis; Chromis = Chromis atripectoralis; Humbug = Dascyllus aruanus; Lemon = Pomacentrus moluccensis; whitedams = Dischistodus perspicillatus
2) Elevated CO2 [CO2] (850-1,050 µatm) or control [Control] (400 - 450 µatm) groups
3) Number of seconds the fish was active per minute, averaged across the duration of the trial
```

Then, let's check the structure of the data to find whether there are some spelling issues in columns `species` and `treatment`.
```{r checkspelling, message = F}
# show the internal structure with str()
str(cdata)
```
Right! We have 6 species and 2 treatment.

Then, let's check whether or not there are some issues in column `activity`. We can plot box plots to see the outliers for each species.
```{r checkactivity, message = F, fig.cap = "Box plots for activity of each species."}
# plot box plot for each species
ggplot(cdata, aes(x = species, y = activity, col = species)) +
  geom_boxplot()
```

It looks great! Although there are some outliers in Figure \@ref(fig:checkactivity), they do not look like mistakes.

Now, we can create a table of summary data that includes: 1) the mean, 2) the standard deviation and 3) the sample sizes of unique fish across all six fish species for each treatment.
```{r summarytab, message = F, tab.cap= "Summary of activity data from Clark et al. (2020)"}
# calculate means, SD and N by species and treatment
cdata_group <- cdata %>% group_by(species, treatment) %>% mutate(size = length(activity), mean_activity = mean(activity), sd_activity = sd(activity))

# delete duplicate data and sort
cdata_summary <- cdata_group %>% .[, c(-3)] %>% .[!duplicated(.), ] %>% .[order(.$species, .$treatment), ]

# use flextable to render the summary table in a tidy format
flextable(cdata_summary) %>% add_footer_lines(., values = c("size, sample size", "mean_activity, mean of activity", "sd_activity, standard deviation of activity")) 
```

## Metadata Formatting

Now, we start to merge the data from Clark et al. (2020) with the "ocean_meta_data.csv".

First, let's re-write the Table \@ref(tab:summarytab) to one species in one row.
```{r widesummarytab, message = F}
# write a new tibble with pivot_wider() to divide the old tibble by treatment
ncdata_summary <- cdata_summary %>% pivot_wider(id_cols = species, names_from = treatment, values_from = c(size, mean_activity, sd_activity))

# have a quick look
ncdata_summary
```

Looks great! Now, let's give the full species name, re-order and re-name these columns to meet format of the metadata file.
```{r resummarytab, message = F, tab.cap= "Wider summary of activity data from Clark et al. (2020)"}
# give the full species name
ncdata_summary$species <- c("Acanthochromis polyacanthus", "Pomacentrus amboinensis", "Chromis atripectoralis", "Dascyllus aruanus", "Pomacentrus moluccensis", "Dischistodus perspicillatus")

# re-order the tibble
ncdata_summary <- ncdata_summary[, c(1, 3, 5, 7, 2, 4, 6)]

# re-name the tibble
names(ncdata_summary) <- c("Species", "ctrl.n", "ctrl.mean", "ctrl.sd", "oa.n", "oa.mean", "oa.sd")

# have a quick look with flextable()
flextable(ncdata_summary) %>% add_footer_lines(., values = c("ctrl.n, sample size of control group", "ctrl.mean, mean activity of control group", "ctrl.sd, standard deviation of activity of control group", "oa.n, sample size of treatment group", "oa.mean, mean activity of treatment group", "oa.sd, standard deviation of activity of treatment group"))
```


Then, we can write the summary statistics in the Table \@ref(tab:resummarytab) into the metadata format.
```{r mergeClark, message = F}
# because we have 6 species and 2 treatments, we need 12 rows to store the summary data
cmdata <- cmrdata
cmdata[1:12, ] <- cmrdata[1, ]

# bind the new tibble
fcdata <- cbind(cmdata, ncdata_summary)
```

Finally, let's add the final tibble into the "ocean_meta_data.csv".
```{r mergefinal, message = F}
# add the tibble with marge() to let the column names to automatch
data <- rbind(mdata, fcdata)

# have a quick look
tail(data, 15)
```

Looks good! But there are still some small problems:

1. column names have spaces. Let's rename them.
```{r renamecol, message = F}
# rename the columns
colnames(data) <- gsub(" ", "_", colnames(data))

# have a quick look
data
```



2. Entries of species names have spaces. Let's rename them again.Less spaces is always better.
```{r renamesp, message = F}
# replace spaces by underline with gsub()
data$Species <- gsub(" ", "_", data$Species)

# have a quick look
head(data)
```

3. There are some "-" in column `Pub_year_IF`, `2017_IF` and `Cue/stimulus_type`. Let's replace them by `NA`.
```{r replace-, message = F}
# replace - by NA with gsub()
data$Pub_year_IF <- gsub("-", NA, data$Pub_year_IF)
data$`2017_IF` <- gsub("-", NA, data$`2017_IF`)
data$`Cue/stimulus_type` <- gsub("-", NA, data$`Cue/stimulus_type`)
```

4. Add a column `observation` to mark each row.
```{r obs, message = F}
# add a column of row numbers
data <- tibble(Obs = 1 : nrow(data), data)

# have a quick look
head(data)
```


Good job! Now, we finish our data cleansing and can output the new metadata file.
```{r outputdata, message = F}
# output a .csv file
write_csv(data, "./out_data/metadata.csv")
```

# Meta-analysis

## Meta-analytis models

First, let's calculate the effect sizes. We will use the log response ratio, which is ‘ROM’ in Metafor.
```{r, RRcal, message = F}
# calculate log response ratio (ROM), where should be given the mean, sample size and standard deviation of experiment and control groups.
InRR <- escalc(measure= "ROM", m1i = oa.mean, sd1i = oa.sd, n1i = oa.n, m2i = ctrl.mean, sd2i = ctrl.sd, n2i = ctrl.n, data = data, var.names = c("InRR", "InRR_V"))

# have a quick look
head(InRR)
```

Oops! We have some NaNs. They are meaningless to our the rest study. Let's remove them.
```{r nanrm, message = F}
# remove rows containing NA
cInRR <- InRR %>% filter(., complete.cases(.))
```

It time to out put our meta-analysis data. In which the column `InRR` is log response ratio, and column `InRR_V` is sampling variance of lnRR. 
```{r outputmetadata, message = F}
# output a .csv file
write_csv(data, "./out_data/InRR_metadata.csv")
```

Now, we start to use a multilevel meta-analytic model.
```{r MLMA, message = F, warning = F}
# do a multilevel meta-analytic model with random effect of study and observation
MLMA <- rma.mv(yi = InRR ~ 1, V = InRR_V,
               # via restricted maximum likelihood
               method = "REML",
               # estimate a random effect variance between study and observation
               random = list(~1 | Study,
                             ~1 | Obs),
               # use t-test rather than z-test
               test = "t",
               # calculate the degrees of freedom based on the lowest clustering level
               dfs = "contain",
               data = cInRR)
MLMA
```

We can see from above: 1) Overall meta-analytic mean is `r coef(MLMA)`, and 2) 95% confidence interval is `r MLMA$ci.lb` to `r MLMA$ci.ub`

## Heterogeneity Analysis

Now, we start to measure the heterogeneity in effect size estimates across studies.
```{r I2, message = F}
# calculate I^2 by i2_ml()
i2_ml(MLMA, data = cInRR)
```

Then, we calculate the prediction interval for heterogeneity
```{r pi, message = F}
# calculate prediction interval by predict()
predict(MLMA)
```

















